Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-07 WSSSPE1_report/WSSSPE1_summary_report.tex
finished first draft of cross-cutting section

In his keynote, Phil Bourne suggested that perhaps sustainability can be defined
as the effort that happens to make the essential things continue. This leads to
having to decide what it is that we want to sustain, if what we want to sustain is
valuable, and finally, who would care if it went away, and how does one measure
how much they care.

@@

Lenhardt~\cite{lenhardt-wssspe1-panel} summarized the contributions of the Modeling Sustainability papers as in Table~\ref{tab:defining-sustainability}. \todo{discussion of this?}

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from \cite{lenhardt-wssspe1-panel}}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
		\hline
{\bf Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
		\hline
Calero, et al. \cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
		\hline
Venters, et al. \cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
		\hline
Pierce, et al. \cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
		\hline
Katz and Proctor \cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
		\hline
Lenhardt, et al. \cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
		\hline
Weber, et al. \cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
		\hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

\todo{Also see discussion at:} \url{https://docs.google.com/document/d/1eVfioGNlihXG_1Y8BgdCI6tXZKrybZgz5XuQHjT1oKU/edit?pli=1#bookmark=id.2pxw5yey8jl0}

>>>>

In the opening keynote, Phil Bourne suggested that perhaps sustainability can be defined
as the effort that happens to make the essential things continue. This leads to
having to decide what it is that we want to sustain, if what we want to sustain is
valuable, and finally, who would care if it went away, and how does one measure
how much they care.

@@

Lenhardt~\cite{lenhardt-wssspe1-panel} summarized the contributions of the Modeling
Sustainability papers in the panel.
As shown in Table~\ref{tab:defining-sustainability}, for each paper he
discussed what was meant by software, if there was a definition of sustainability, and
what the approach was to either understand or evaluate sustainability.

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from \cite{lenhardt-wssspe1-panel}}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
		\hline
{\bf Paper/Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
		\hline
Calero, et al. \cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
		\hline
Venters, et al. \cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
		\hline
Pierce, et al. \cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
		\hline
Katz and Proctor \cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
		\hline
Lenhardt, et al. \cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
		\hline
Weber, et al. \cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
		\hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

Finally, in the workshop's closing session, one of the discussion topics was 
what success would look like for the set of WSSSPE activities beyond just the workshop.
One of the answers that was suggested was
having an agreed version of what we mean by sustainability.
